The Stakes of Scrying: \lstinline[style=inlinecode]{.^} dotket and the Seer Proposal

### 2024-01-23 mss.tex
Post new style file

%  Make first page footer:
\fancypagestyle{firststyle}{%
\fancyhf{}% Clear header/footer
\fancyhead{}
\fancyfoot[L]{{\footnotesize
              %% We toggle between these:
              Manuscript submitted for review.\\
              % {\it Urbit Systems Technical Journal} I:1 (2024):  1–3. \\
              ~ \\
              Address author correspondence to \patp{rovnys-ricfer}.
              }}
}
%  Arrange subsequent pages:
\fancyhf{}
\fancyhead[LE]{{\urbitfont Urbit Systems Technical Journal}}
\fancyhead[RO]{The Stakes of Scrying:  \texttt{\string.\string^} dotket and the Seer Proposal}
\fancyfoot[LE,RO]{\thepage}

\newcommand{\dotket}{\texttt{\string.\string^}~dotket}

%%MANUSCRIPT
\title{The Stakes of Scrying:  \texttt{\string.\string^} dotket and the Seer Proposal}
\author{\authorname~\authorpatp \\ \affiliation}
\date{}

@@

The Nock 12 opcode and the \dotket~rune are used by userspace code to ask the kernel for data in the scry namespace.\footnote{The scry namespace, in brief, is a collection of immutable values “bound” by convention at fixed endpoints.  The scry namespace is defined by a unique path for each resource.  A representative scry path would have the entries \texttt{/host/rift/life/vane/request-type/revision-number/desk/file-path/mark}.}  When Urbit runs userspace code, that code is not raw Nock.  It is instead called Mock, which is Nock with two additions:   deterministic errors are caught and returned as values to the kernel; and there is an extra Nock opcode, opcode 12.  A simple Nock 12 formula looks like \texttt{\string[12 \string[1 /foo/bar/baz\string]\string]}, where \texttt{/foo/bar/baz} is a path literal \texttt{\string[\%foo \%bar \%baz \textasciitilde\string]}, representing a path in the scry namespace.  The result of running this opcode will be the result of performing the scry request, i.e. dereferencing the path.

Catching deterministic errors is uncontroversial, but the Nock 12 opcode poses some difficulties and has been under scrutiny for years now.  The most tangible problem with Nock 12 is that it prevents userspace code from persistently memoizing Nock computations; at the moment, only kernelspace code can hold onto Nock memoization caches across multiple Arvo events.  The lifetime of a userspace memoization cache is limited to a single metacircular evaluation context, i.e. a single call to \texttt{++mink} or an equivalent gate that runs Mock (virtual Nock with the extra opcode).

Userspace code cannot memoize Nock persistently because it is not feasible for the runtime to prevent a jet mismatch in the metacircular evaluator gate.  A gate like \texttt{++mink} takes in a Mock expression (a cell of \texttt{\string[subject formula\string]}) and a “scry handler gate” (which is described shortly), and returns a value which is one of:

\begin{itemize}
  \item  \texttt{\string[0 computation-result\string]} for a successful evaluation;
  \item  \texttt{\string[1 block\string]} in the case of a failed scry request; or
  \item  \texttt{\string[2 error\string]} in the case of a deterministic error.
\end{itemize}

\noindent
While error handling is not relevant to the current question, a brief description is included here for thoroughness.  The \texttt{error} in \texttt{++mink}'s result \texttt{\string[2 error\string]} is a deterministic error, meaning that every attempt to evaluate that \texttt{\string[subject formula\string]} pair will always result in that same error value—deterministic errors correspond to the lines in the Nock spec where \texttt{*a} evaluates to \texttt{*a}, i.e. a trivial infinite loop.  Nondeterministic errors, such as running out of usable memory or getting interrupted by the user, result in the function failing to complete, and the Arvo event will be aborted and rolled back to its state before trying to run this new event.

As mentioned, the \texttt{\string[subject formula\string]} pair is not the only input to \texttt{++mink}.  There is also an argument of the form \texttt{scry=\$-(\string^ (unit (unit)))}, called the scry handler gate.  When \texttt{++mink} encounters a Nock 12 opcode while evaluating the \texttt{\string[subject formula\string]} expression, it evaluates the sub-formula to obtain a path to look up in the scry namespace.  Once it has this path, it calls the scry handler gate on the path, which returns one of:

\begin{itemize}
  \item  \texttt{\string[\textasciitilde~\textasciitilde~value\string]} positive scry response;
  \item  \texttt{\string[\textasciitilde~\textasciitilde\string] }permanent scry failure; or
  \item  \texttt{\textasciitilde} ``block''.
\end{itemize}

\noindent
The positive response case means that the scry was performed successfully and yielded a value, in which case \texttt{++mink} injects the value as the result of the Nock 12 opcode and continues to evaluate the Nock.  The permanent failure means the scry namespace knows there will never be a value at that path, so \texttt{++mink} can immediately return a deterministic error.

A scry block means the system refused to answer the question, i.e. the request was "blocked" for some reason—this could be due to lack of permission, or because the request asked about the future, or the request asked for data from another ship that isn't synchronously available within this Arvo event. If the scry blocks, \texttt{++mink} returns a \texttt{\string[1 block\string]}, where \texttt{block} is the scry request path whose value could not be resolved.

The Nock 12 opcode is problematic because it renders the result of the computation dependent on external data which are not in the \texttt{\string[subject formula\string]} that is being evaluated.  The result is no longer a pure function of the \texttt{subject} and \texttt{formula}:  depending on what the scry handler gate is, the same \texttt{\string[subject formula\string]} could yield different results.  Consider \texttt{\string[subject formula\string]} of \texttt{\string[0 \string[12 1 /foo/bar\string]\string]}.  This returns the result of looking up the \texttt{/foo/bar} scry request path from the scry gate. Userspace code (which is untrusted) could call \texttt{(mink \string[\string[0 12 1 /foo/bar\string]\string] |=(\string^ \string[\textasciitilde~\textasciitilde~33\string]))}, acquire the result \texttt{33}, then call \texttt{(mink \string[\string[0 12 1 /foo/bar\string]\string] |=(\string^ \string[\textasciitilde~\textasciitilde~44\string]))} to yield \texttt{44}.  The \texttt{subject}s and \texttt{formula}s were identical, but because they were evaluated with different scry gates, they yielded different results.

@@

As long as a correct Arvo kernel is supplying the scry handler, Mock retains its purely functional nature.  But consider again the case where userspace code runs the same \texttt{\string[subject formula\string]} twice, but with two different scry gates that give different answers to some scry path. From the perspective of a simple Nock interpreter, it's still deterministic:  \texttt{++mink} is just some Nock function that runs some other computation (Mock) on some inputs.  Because it's just some Nock function, it can't break determinism of the Arvo event; replaying the Arvo event will still yield the same result.  The problem only arises when trying to get clever about how the runtime evaluates Nock, in particular by caching the results of Nock computations.

Development versions of Vere support persistent Nock memoization.  This means a programmer can emit a Nock hint to the interpreter asking the interpreter to cache the result of the computation, so that if it gets run again, the result will be retrieved from the runtime's cache instead of rerun step by step.  The cache key is a \texttt{\string[subject formula\string]} pair.  When the interpreter sees a \texttt{\%memo} hint, it looks at the \texttt{\string[subject formula\string]} that the hint surrounds. If the memoization cache has a value for that key, the runtime uses that value as the result of the computation.  Otherwise the runtime runs the computation, inserts the result into the cache, and returns the result.

@@

\begin{enumerate}
  \item  Let userspace code place arbitrary values into the cache.  This poses a severe security problem:  event replay would be nondeterministic, influenced by the contents of the memoization cache, allowing malicious userspace agents to perform a kind of denial of service attack by preventing event replay, or possibly even worse, by injecting cache lines that other applications would use, causing those other applications to malfunction in a way the malicious code wants.  The worst case would be causing a system app such as \texttt{\%hood} to fail to enforce a security check, allowing the malicious code to escalate its own privileges and permitting it to ask the system app to do something like reinstall the kernel.

  \item  Place the scry gate into the cache key for the memoization cache, so a cache key would be \texttt{\string[\string[subject formula\string] scry-gate\string]} instead of just \texttt{\string[subject formula\string]}.  This would close the security hole, but since Arvo's scry gate undergoes small changes on between every agent activation, it would make these cache lines useless outside of the immediate execution context within a single \texttt{++mink} call.  This is exactly the situation userspace code is already in, and thus not helpful at all.

  \item  Have the interpreter record the set of scry \texttt{\string[path value\string]} pairs looked up by the computation.  This would allow userspace code to perform scry requests safely, but the cost is a nontrivial amount of machinery in the runtime, along with a computational cost of looking up a cache line that grows linearly with the number of scry paths that were requested.  Of the three options listed here, this is the most reasonable, but it is still highly nontrivial and the requirement to compare all requested paths could be onerous.
\end{enumerate}

\noindent
Instead of implementing any of these options, Urbit has currently simple disabled the ability for userspace code to modify the Nock memoization cache.  This is enforced by the jet for \texttt{++mink}; the jet is simply a call back into the normal Nock interpreter, which is really a Mock interpreter, whose state includes a list of nouns representing the stack of scry handler gates for the current execution context.  If this list is null \texttt{\textasciitilde}, then we must be in the kernel and the code is allowed to modify the memoization cache; otherwise, we are in some kind of Mock context, and the code is not allowed to modify the cache.

@@

Seer is a monadic scry interface which allows application developers to read from the namespace without using \dotket~or its virtual Nock operator 12.  Userspace code run inside of this gate can reestablish the guarantee that it is purely functional, allowing it to modify the memoization cache.\footnote{Note that it's still acceptable for that gate to be called from Mock code, or even nested Mock code, that had previously performed scry requests that were not referentially transparent:  raw Nock is context-free, so no matter how the \texttt{\string[subject formula\string]} cell was created, the result of its evaluation is still pure.}

@@

\begin{lstlisting}
++  on-poke
  |=  [m=mark v=vase]
  ^-  (quip card _this)
  =/  foo-result  .^  @da  /cx/(now)/foo
  ::  do something with foo-result...
  [~ this]
\end{lstlisting}

A corresponding snippet of a Gall agent that scries using \texttt{++seer} in the continuation style would look like this:

\begin{lstlisting}
++  gear         ::  gall seer
  |*  a=mold
  (seer vase a)
++  on-poke
  |=  [m=mark v=vase]
  ^-  (gear (quip card _this))  ::  (seer vase (quip card _this))
  :+  %scry  /cx/(now)/foo  |=  foo=vase
  =/  foo-result  !<(@ foo)
  ::  do something with foo-result...
  [%done ~ this]
\end{lstlisting}

\texttt{++seer} values can be composed using a monadic bind:

\begin{lstlisting}
++  foo  ^-  (seer vase @)  !!  ::  /foo
++  bar  ^-  (seer vase @)  !!  ::  /bar
++  on-poke
  |=  [m=mark v=vase]
  =*  r  (quip card _this)
  ^-  (gear r)
  =/  bind  (rapt vase r)
  ;<  foo=@  bind  foo
  ;<  bar=@  bind  bar
  [%done ~ this(some-state (add foo bar))]
\end{lstlisting}

\noindent
(In this example, use is made of \texttt{++rapt}.  \texttt{++rapt} is a monadic scry binding gate \citep{Levan2023b}.)

Here, a \texttt{++seer} is the type of programs that scry:

\begin{lstlisting}
++  seer
  |*  [r=mold a=mold]
  $~  [%done *a]
  $%  [%done p=a]
      [%scry p=path k=$-(r (seer r a))]
  ==
\end{lstlisting}

The \texttt{++seer} scry system is intentionally incompatible with the current \dotket~scry system.  We believe that \texttt{++seer} degrades ergonomics only slightly.  In the initial introduction, we shall provide a wrapper library which accepts an agent written in the \texttt{++seer} style and produces an agent that scries in the \dotket~style.

@@

A Gall agent will be able to emit a new kind of move, in addition to all the moves it can emit now, that asks Gall to perform a scry request on the agent's behalf.  When Gall processes this move, it will perform the scry request and convert the result of the scry into a response move that will be delivered back into the agent in a later activation of the Gall vane, in accordance with the normal move processing order that Gall and Arvo use.  Thus an agent could emit a list of moves such as \texttt{\string[poke-1 scry-A poke-2 scry-B poke-3 \textasciitilde\string]} and Gall will handle all of those moves in the same order as if they were all pokes, i.e. the kernel will not introduce any exceptions to its normal move processing order for scry request moves.

@@

Gall will also need to run the \texttt{++on-peek} arm of a Gall agent in the Seer monad, i.e. that arm will return either \texttt{\string[\%done scry-result\string]} to indicate it's done or \texttt{\string[\%scry scry-request-path callback\string]} to ask Gall to perform a scry request, in which case Gall performs the scry, injects the result into the callback gate, and repeats, until a \texttt{\%done} is returned.  This should not be a direct problem, but betokens complexity creeping into the Gall interface.

>>>>

%  Make first page footer:
\fancypagestyle{firststyle}{%
\fancyhf{}% Clear header/footer
\fancyhead{}
\fancyfoot[L]{{\footnotesize
              %% We toggle between these:
              Manuscript submitted for review.\\
              % {\it Urbit Systems Technical Journal} I:1 (2024):  1–3. \\
              ~ \\
              Address author correspondence to \patp{rovnys-ricfer}.
              }}
}
%  Arrange subsequent pages:
\fancyhf{}
\fancyhead[LE]{{\urbitfont Urbit Systems Technical Journal}}
\fancyhead[RO]{The Stakes of Scrying:  \lstinline[style=inlinecode]{.^} dotket and the Seer Proposal}
\fancyfoot[LE,RO]{\thepage}

\newcommand{\dotket}{\lstinline[style=inlinecode]{.^}~dotket}

%%MANUSCRIPT
\title{The Stakes of Scrying:  \lstinline[style=inlinecode]{.^} dotket and the Seer Proposal}
\author{\authorname~\authorpatp \\ \affiliation}
\date{}

@@

The Nock 12 opcode and the \dotket~rune are used by userspace code to ask the kernel for data in the scry namespace.\footnote{The scry namespace, in brief, is a collection of immutable values “bound” by convention at fixed endpoints.  The scry namespace is defined by a unique path for each resource.  A representative scry path would have the entries \lstinline[style=inlinecode]{/host/rift/life/vane/request-type/revision-number/desk/file-path/mark}.}  When Urbit runs userspace code, that code is not raw Nock.  It is instead called Mock, which is Nock with two additions:   deterministic errors are caught and returned as values to the kernel; and there is an extra Nock opcode, opcode 12.  A simple Nock 12 formula looks like \lstinline[style=inlinecode]{[12 [1 /foo/bar/baz]]}, where \lstinline[style=inlinecode]{/foo/bar/baz} is a path literal \lstinline[style=inlinecode]{[%foo %bar %baz ~]}, representing a path in the scry namespace.  The result of running this opcode will be the result of performing the scry request, i.e. dereferencing the path.

Catching deterministic errors is uncontroversial, but the Nock 12 opcode poses some difficulties and has been under scrutiny for years now.  The most tangible problem with Nock 12 is that it prevents userspace code from persistently memoizing Nock computations; at the moment, only kernelspace code can hold onto Nock memoization caches across multiple Arvo events.  The lifetime of a userspace memoization cache is limited to a single metacircular evaluation context, i.e. a single call to \lstinline[style=inlinecode]{++mink} or an equivalent gate that runs Mock (virtual Nock with the extra opcode).

Userspace code cannot memoize Nock persistently because it is not feasible for the runtime to prevent a jet mismatch in the metacircular evaluator gate.  A gate like \lstinline[style=inlinecode]{++mink} takes in a Mock expression (a cell of \lstinline[style=inlinecode]{[subject formula]}) and a “scry handler gate” (which is described shortly), and returns a value which is one of:

\begin{itemize}
  \item  \lstinline[style=inlinecode]{[0 computation-result]} for a successful evaluation;
  \item  \lstinline[style=inlinecode]{[1 block]} in the case of a failed scry request; or
  \item  \lstinline[style=inlinecode]{[2 error]} in the case of a deterministic error.
\end{itemize}

\noindent
While error handling is not relevant to the current question, a brief description is included here for thoroughness.  The \lstinline[style=inlinecode]{error} in \lstinline[style=inlinecode]{++mink}'s result \lstinline[style=inlinecode]{[2 error]} is a deterministic error, meaning that every attempt to evaluate that \lstinline[style=inlinecode]{[subject formula]} pair will always result in that same error value—deterministic errors correspond to the lines in the Nock spec where \lstinline[style=inlinecode]{*a} evaluates to \lstinline[style=inlinecode]{*a}, i.e. a trivial infinite loop.  Nondeterministic errors, such as running out of usable memory or getting interrupted by the user, result in the function failing to complete, and the Arvo event will be aborted and rolled back to its state before trying to run this new event.

As mentioned, the \lstinline[style=inlinecode]{[subject formula]} pair is not the only input to \lstinline[style=inlinecode]{++mink}.  There is also an argument of the form \lstinline[style=inlinecode]{scry=$-(^ (unit (unit)))}, called the scry handler gate.  When \lstinline[style=inlinecode]{++mink} encounters a Nock 12 opcode while evaluating the \lstinline[style=inlinecode]{[subject formula]} expression, it evaluates the sub-formula to obtain a path to look up in the scry namespace.  Once it has this path, it calls the scry handler gate on the path, which returns one of:

\begin{itemize}
  \item  \lstinline[style=inlinecode]{[~ ~ value]} positive scry response;
  \item  \lstinline[style=inlinecode]{[~ ~] }permanent scry failure; or
  \item  \lstinline[style=inlinecode]{~} ``block''.
\end{itemize}

\noindent
The positive response case means that the scry was performed successfully and yielded a value, in which case \lstinline[style=inlinecode]{++mink} injects the value as the result of the Nock 12 opcode and continues to evaluate the Nock.  The permanent failure means the scry namespace knows there will never be a value at that path, so \lstinline[style=inlinecode]{++mink} can immediately return a deterministic error.

A scry block means the system refused to answer the question, i.e. the request was "blocked" for some reason—this could be due to lack of permission, or because the request asked about the future, or the request asked for data from another ship that isn't synchronously available within this Arvo event. If the scry blocks, \lstinline[style=inlinecode]{++mink} returns a \lstinline[style=inlinecode]{[1 block]}, where \lstinline[style=inlinecode]{block} is the scry request path whose value could not be resolved.

The Nock 12 opcode is problematic because it renders the result of the computation dependent on external data which are not in the \lstinline[style=inlinecode]{[subject formula]} that is being evaluated.  The result is no longer a pure function of the \lstinline[style=inlinecode]{subject} and \lstinline[style=inlinecode]{formula}:  depending on what the scry handler gate is, the same \lstinline[style=inlinecode]{[subject formula]} could yield different results.  Consider \lstinline[style=inlinecode]{[subject formula]} of \lstinline[style=inlinecode]{[0 [12 1 /foo/bar]]}.  This returns the result of looking up the \lstinline[style=inlinecode]{/foo/bar} scry request path from the scry gate. Userspace code (which is untrusted) could call \lstinline[style=inlinecode]{(mink [[0 12 1 /foo/bar]] |=(^ [~ ~ 33]))}, acquire the result \lstinline[style=inlinecode]{33}, then call \lstinline[style=inlinecode]{(mink [[0 12 1 /foo/bar]] |=(^ [~ ~ 44]))} to yield \lstinline[style=inlinecode]{44}.  The \lstinline[style=inlinecode]{subject}s and \lstinline[style=inlinecode]{formula}s were identical, but because they were evaluated with different scry gates, they yielded different results.

@@

As long as a correct Arvo kernel is supplying the scry handler, Mock retains its purely functional nature.  But consider again the case where userspace code runs the same \lstinline[style=inlinecode]{[subject formula]} twice, but with two different scry gates that give different answers to some scry path. From the perspective of a simple Nock interpreter, it's still deterministic:  \lstinline[style=inlinecode]{++mink} is just some Nock function that runs some other computation (Mock) on some inputs.  Because it's just some Nock function, it can't break determinism of the Arvo event; replaying the Arvo event will still yield the same result.  The problem only arises when trying to get clever about how the runtime evaluates Nock, in particular by caching the results of Nock computations.

Development versions of Vere support persistent Nock memoization.  This means a programmer can emit a Nock hint to the interpreter asking the interpreter to cache the result of the computation, so that if it gets run again, the result will be retrieved from the runtime's cache instead of rerun step by step.  The cache key is a \lstinline[style=inlinecode]{[subject formula]} pair.  When the interpreter sees a \lstinline[style=inlinecode]{%memo} hint, it looks at the \lstinline[style=inlinecode]{[subject formula]} that the hint surrounds. If the memoization cache has a value for that key, the runtime uses that value as the result of the computation.  Otherwise the runtime runs the computation, inserts the result into the cache, and returns the result.

@@

\begin{enumerate}
  \item  Let userspace code place arbitrary values into the cache.  This poses a severe security problem:  event replay would be nondeterministic, influenced by the contents of the memoization cache, allowing malicious userspace agents to perform a kind of denial of service attack by preventing event replay, or possibly even worse, by injecting cache lines that other applications would use, causing those other applications to malfunction in a way the malicious code wants.  The worst case would be causing a system app such as \lstinline[style=inlinecode]{%hood} to fail to enforce a security check, allowing the malicious code to escalate its own privileges and permitting it to ask the system app to do something like reinstall the kernel.

  \item  Place the scry gate into the cache key for the memoization cache, so a cache key would be \lstinline[style=inlinecode]{[[subject formula] scry-gate]} instead of just \lstinline[style=inlinecode]{[subject formula]}.  This would close the security hole, but since Arvo's scry gate undergoes small changes on between every agent activation, it would make these cache lines useless outside of the immediate execution context within a single \lstinline[style=inlinecode]{++mink} call.  This is exactly the situation userspace code is already in, and thus not helpful at all.

  \item  Have the interpreter record the set of scry \lstinline[style=inlinecode]{[path value]} pairs looked up by the computation.  This would allow userspace code to perform scry requests safely, but the cost is a nontrivial amount of machinery in the runtime, along with a computational cost of looking up a cache line that grows linearly with the number of scry paths that were requested.  Of the three options listed here, this is the most reasonable, but it is still highly nontrivial and the requirement to compare all requested paths could be onerous.
\end{enumerate}

\noindent
Instead of implementing any of these options, Urbit has currently simple disabled the ability for userspace code to modify the Nock memoization cache.  This is enforced by the jet for \lstinline[style=inlinecode]{++mink}; the jet is simply a call back into the normal Nock interpreter, which is really a Mock interpreter, whose state includes a list of nouns representing the stack of scry handler gates for the current execution context.  If this list is null \lstinline[style=inlinecode]{~}, then we must be in the kernel and the code is allowed to modify the memoization cache; otherwise, we are in some kind of Mock context, and the code is not allowed to modify the cache.

@@

Seer is a monadic scry interface which allows application developers to read from the namespace without using \dotket~or its virtual Nock operator 12.  Userspace code run inside of this gate can reestablish the guarantee that it is purely functional, allowing it to modify the memoization cache.\footnote{Note that it's still acceptable for that gate to be called from Mock code, or even nested Mock code, that had previously performed scry requests that were not referentially transparent:  raw Nock is context-free, so no matter how the \lstinline[style=inlinecode]{[subject formula]} cell was created, the result of its evaluation is still pure.}

@@

\begin{lstlisting}[style=listingcode]
++  on-poke
  |=  [m=mark v=vase]
  ^-  (quip card _this)
  =/  foo-result  .^  @da  /cx/(now)/foo
  ::  do something with foo-result...
  [~ this]
\end{lstlisting}

A corresponding snippet of a Gall agent that scries using \lstinline[style=inlinecode]{++seer} in the continuation style would look like this:

\begin{lstlisting}[style=listingcode]
++  gear         ::  gall seer
  |*  a=mold
  (seer vase a)
++  on-poke
  |=  [m=mark v=vase]
  ^-  (gear (quip card _this))  ::  (seer vase (quip card _this))
  :+  %scry  /cx/(now)/foo  |=  foo=vase
  =/  foo-result  !<(@ foo)
  ::  do something with foo-result...
  [%done ~ this]
\end{lstlisting}

\lstinline[style=inlinecode]{++seer} values can be composed using a monadic bind:

\begin{lstlisting}[style=listingcode]
++  foo  ^-  (seer vase @)  !!  ::  /foo
++  bar  ^-  (seer vase @)  !!  ::  /bar
++  on-poke
  |=  [m=mark v=vase]
  =*  r  (quip card _this)
  ^-  (gear r)
  =/  bind  (rapt vase r)
  ;<  foo=@  bind  foo
  ;<  bar=@  bind  bar
  [%done ~ this(some-state (add foo bar))]
\end{lstlisting}

\noindent
(In this example, use is made of \lstinline[style=inlinecode]{++rapt}.  \lstinline[style=inlinecode]{++rapt} is a monadic scry binding gate \citep{Levan2023b}.)

Here, a \lstinline[style=inlinecode]{++seer} is the type of programs that scry:

\begin{lstlisting}[style=listingcode]
++  seer
  |*  [r=mold a=mold]
  $~  [%done *a]
  $%  [%done p=a]
      [%scry p=path k=$-(r (seer r a))]
  ==
\end{lstlisting}

The \lstinline[style=inlinecode]{++seer} scry system is intentionally incompatible with the current \dotket~scry system.  We believe that \lstinline[style=inlinecode]{++seer} degrades ergonomics only slightly.  In the initial introduction, we shall provide a wrapper library which accepts an agent written in the \lstinline[style=inlinecode]{++seer} style and produces an agent that scries in the \dotket~style.

@@

A Gall agent will be able to emit a new kind of move, in addition to all the moves it can emit now, that asks Gall to perform a scry request on the agent's behalf.  When Gall processes this move, it will perform the scry request and convert the result of the scry into a response move that will be delivered back into the agent in a later activation of the Gall vane, in accordance with the normal move processing order that Gall and Arvo use.  Thus an agent could emit a list of moves such as \lstinline[style=inlinecode]{[poke-1 scry-A poke-2 scry-B poke-3 ~]} and Gall will handle all of those moves in the same order as if they were all pokes, i.e. the kernel will not introduce any exceptions to its normal move processing order for scry request moves.

@@

Gall will also need to run the \lstinline[style=inlinecode]{++on-peek} arm of a Gall agent in the Seer monad, i.e. that arm will return either \lstinline[style=inlinecode]{[%done scry-result]} to indicate it's done or \lstinline[style=inlinecode]{[%scry scry-request-path callback]} to ask Gall to perform a scry request, in which case Gall performs the scry, injects the result into the callback gate, and repeats, until a \lstinline[style=inlinecode]{%done} is returned.  This should not be a direct problem, but betokens complexity creeping into the Gall interface.
